Untitled

### 2018-06-13 MS-cclifeexpec_paa.Rmd
from computer.

<!-- # Methods -->
We estimate changes in life expectancy by combining three primary datasets: excess mortality data due to meterological hazards related to climate change from Forzieri et al. [@forzieri2017increasing], cause-of-death distributions from the Global Burden of Disease study (GBD) [@GBD;@wang2012age], and life tables from the Human Mortality Database (HMD) [@HMD]. Forzieri et al. [@forzieri2017increasing] focused on the projected excess mortality of heatwaves, coldwaves, wildfires, droughts, river and coastal floods, and windstorms under the business-as-usual greenhouse gas emission scenario (SRES A1B). They modelled long-term, small-area demographic changes in European countries that correspond to a middle-of-the-road socioeconomic scenario (SSP2 with medium fertility, medium mortality, medium migration, and the Global Education Trend education scenario) and based their projected mortality on extrapolations of an exhaustive examination of contemporary disaster databases within spatially rectified, high-resolution gridded datasets. While the disaster data are not standardized across country, Forzeri et al. imputed data in incomplete time periods and countries. This imputation could mask spatial variability at the sub-national level, but here we use their country-aggregated results. Their results represent business-as-usual climate change and human development without incorporating potential adaptation. They found a potential 150,000+ climate change related fatalities per year by the mid 2080s with climate change contributing to 90% of mortality rise as opposed to population changes. These data provide the anticipated fatalities but not the effect these deaths might have on life expectancy.

@@

<!-- # Results -->
We find that climate change could alter life expectancy by `r round(mean(lt_lancet$DIF_MID),2)` years (-2.5 months (`r round(mean(lt_lancet$DIF_LOW),2)` to `r round(mean(lt_lancet$DIF_HIGH),2)` years)) in the average European country (**\autoref{figure1}**). This reduction is comparable to mortality due to influenza and pneumonia [@arias2013united] in the United States. Although the average European country could see a change of `r round(mean(lt_lancet$DIF_MID),2)` years, several countries are likely to experience considerably greater reductions in life expectancy. Luxembourg could see a reduction of up to `r round(lt_lancet$DIF_HIGH[which(lt_lancet$CNTRY == "LUX")],2)*-1` years and the medium-variant of climate hazards could cost the average Spaniard `r round(lt_lancet$DIF_MID[which(lt_lancet$CNTRY == "ESP")],2)*-1` years of life.

@@

<!-- # Results -->
We find that climate change could alter life expectancy by `r round(mean(lt_lancet$DIF_MID),2)` years (-2.5 months (`r round(mean(lt_lancet$DIF_LOW),2)` - `r round(mean(lt_lancet$DIF_HIGH),2)` years)) in the average European country (**Fig. 1**). Although the average European country could see a change of -0.22 years, several countries are likely to experience considerably greater reductions in life expectancy. Luxembourg could see a change of up to -2.2 years and the medium-variant of climate hazards could cost the average Spaniard 0.98 years of life.

@@

@@

>>>>

# Methods
We estimate changes in life expectancy by combining three primary datasets: excess mortality data due to meterological hazards related to climate change from Forzieri et al. [@forzieri2017increasing], cause-of-death distributions from the Global Burden of Disease study (GBD) [@GBD;@wang2012age], and life tables from the Human Mortality Database (HMD) [@HMD]. Forzieri et al. [@forzieri2017increasing] focused on the projected excess mortality of heatwaves, coldwaves, wildfires, droughts, river and coastal floods, and windstorms under the business-as-usual greenhouse gas emission scenario (SRES A1B). They modelled long-term, small-area demographic changes in European countries that correspond to a middle-of-the-road socioeconomic scenario (SSP2 with medium fertility, medium mortality, medium migration, and the Global Education Trend education scenario) and based their projected mortality on extrapolations of an exhaustive examination of contemporary disaster databases within spatially rectified, high-resolution gridded datasets. While the disaster data are not standardized across country, Forzeri et al. imputed data in incomplete time periods and countries. This imputation could mask spatial variability at the sub-national level, but here we use their country-aggregated results. Their results represent business-as-usual climate change and human development without incorporating potential adaptation. They found a potential 150,000+ climate change related fatalities per year by the mid 2080s with climate change contributing to 90% of mortality rise as opposed to population changes. These data provide the anticipated fatalities but not the effect these deaths might have on life expectancy.

@@

# Results
We find that climate change could alter life expectancy by `r round(mean(lt_lancet$DIF_MID),2)` years (-2.5 months (`r round(mean(lt_lancet$DIF_LOW),2)` to `r round(mean(lt_lancet$DIF_HIGH),2)` years)) in the average European country (**\autoref{figure1}**). This reduction is comparable to mortality due to influenza and pneumonia [@arias2013united] in the United States. Although the average European country could see a change of `r round(mean(lt_lancet$DIF_MID),2)` years, several countries are likely to experience considerably greater reductions in life expectancy. Luxembourg could see a reduction of up to `r round(lt_lancet$DIF_HIGH[which(lt_lancet$CNTRY == "LUX")],2)*-1` years and the medium-variant of climate hazards could cost the average Spaniard `r round(lt_lancet$DIF_MID[which(lt_lancet$CNTRY == "ESP")],2)*-1` years of life.

@@

# Results
We find that climate change could alter life expectancy by `r round(mean(lt_lancet$DIF_MID),2)` years (-2.5 months (`r round(mean(lt_lancet$DIF_LOW),2)` - `r round(mean(lt_lancet$DIF_HIGH),2)` years)) in the average European country (**Fig. 1**). Although the average European country could see a change of -0.22 years, several countries are likely to experience considerably greater reductions in life expectancy. Luxembourg could see a change of up to -2.2 years and the medium-variant of climate hazards could cost the average Spaniard 0.98 years of life.

@@

```{r echo = FALSE, fig.cap=paste("Change in life expectancy at birth ($e_0$) due to business-as-usual climate change. We report changes in life expectancies due to climate change for twenty-eight European countries. The central values represent the ensemble median while the stems represent the upper and lower bounds of the inter-model climate variability.")}

@@

```{r echo = FALSE, warning= FALSE, message = FALSE, fig.cap=paste("Estimated reduction in life expectancy at birth ($e_0$) by the 2080s under the $MID$ scenario.")}

### 2018-06-13 MS-cclifeexpec_paa.tex
from computer.

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_files/figure-latex/figure1-1.pdf}
\caption{Change in life expectancy at birth (\(e_0\)) due to
business-as-usual climate change. We report changes in life expectancies
due to climate change for twenty-eight European countries. The central
values represent the ensemble median while the stems represent the upper
and lower bounds of the inter-model climate variability.\label{figure1}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_files/figure-latex/figure2-1.pdf}
\caption{Estimated reduction in life expectancy at birth (\(e_0\)) by
the 2080s under the \(MID\) scenario.\label{map}}
\end{figure}

@@

@@

@@

@@

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_files/figure-latex/suppfig-1.pdf}
\caption{\textbf{Mortality distribution of Heat-related mortality in 28
European countries}. The thick black line is the average\label{suppfig}}
\end{figure}

\subsection{Methods}\label{methods-1}

>>>>

\section{Results}\label{results}

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_paa_files/figure-latex/figure1-1.pdf}
\caption{Change in life expectancy at birth (\(e_0\)) due to
business-as-usual climate change. We report changes in life expectancies
due to climate change for twenty-eight European countries. The central
values represent the ensemble median while the stems represent the upper
and lower bounds of the inter-model climate variability.\label{figure1}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_paa_files/figure-latex/figure2-1.pdf}
\caption{Estimated reduction in life expectancy at birth (\(e_0\)) by
the 2080s under the \(MID\) scenario.\label{map}}
\end{figure}

@@

\section{Results}\label{results-1}

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_paa_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Change in life expectancy at birth (\(e_0\)) due to
business-as-usual climate change. We report changes in life expectancies
due to climate change for twenty-eight European countries. The central
values represent the ensemble median while the stems represent the upper
and lower bounds of the inter-model climate variability.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_paa_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Estimated reduction in life expectancy at birth (\(e_0\)) by
the 2080s under the \(MID\) scenario.}
\end{figure}

@@

\section{Methods}\label{methods-1}

@@

\begin{figure}
\centering
\includegraphics{MS-cclifeexpec_paa_files/figure-latex/suppfig-1.pdf}
\caption{\textbf{Mortality distribution of Heat-related mortality in 28
European countries}. The thick black line is the average\label{suppfig}}
\end{figure}

\subsection{Methods}\label{methods-2}

### 2018-06-13 PRESENTATIONS/presentation.Rmd
from computer.

---
title: "Climate change projected to reduce European life expectancy"
author: | 
 | Mathew E. Hauer^1^ & Alexis R. Santos^2^
 | ^1^University of Georgia | ^2^Pennsylvania State University
date: "April 26, 2018"
output:   beamer_presentation
classoption: "aspectratio=169"

---

@@

**Mathew E Hauer**
Applied Demographer 
Carl Vinson Institute of Government \
hauer@uga.edu

**Alexis R Santos **
Applied Demographer
Pennsylvania State University

>>>>

---
title: Climate Change Projected to Reduce European Life Expectancy
# subtitle: (A Subtitle Would Go Here if This Were a Class)
author: | 
 | Mathew E. Hauer^1^ & Alexis R. Santos^2^
institute:  | 
 | ^1^University of Georgia | ^2^Pennsylvania State University
titlegraphic1: logo-uga-color.eps
titlegraphic2: pennstate.png
fontsize: 10pt
output:
 beamer_presentation:
    template: ../PRESENTATIONS/presentation_template.tex
    keep_tex: TRUE
make149: true
---

@@

**Mathew E Hauer**\newline
Applied Demographer \newline
Carl Vinson Institute of Government \newline
hauer@uga.edu
\newline
\newline

**Alexis R Santos **\newline
Applied Demographer\newline
Pennsylvania State University

### 2018-06-13 SCRIPTS/001-ANALYSIS_md.Rmd
from computer.

---
title: "Climate Change Projected to Reduce European Life Expectancy"
output:
  word_document: default
  pdf_document: default
date: "April 5, 2018"
---

@@

## R Data Load and Processing The following code chunk will 
1. Import the underlying Forzieri et al data. 
2. Import the Standard European Population from EuroStat. 
3. Import the leading causes of death with age-standardized death rates from EuroStat. 
4. Import the Global Burden of Disease data. This data is located on github. 
5. Download all of the Human Mortality Database data.

To successfully run this code, several R pacakgse must be installed and these
packages are located just below this text section but are tidyverse, HMHFDplus,
tmap, tmaptools, RColorBrewer, data.table, and getPass.

To successfully run this code, you must also have an account at the HMD and
input your username and password when prompted.

@@

# Figure 1
This chunk of code reproduces Figure 1.
```{r Figure1}

@@

# Figure 2
This chunk of code reproduecs Figure 2.
```{r Figure2}

@@

# Supplementay Figure 1
This chunk of code reproduces the Supplementary Figure
```{r suppfig}
GBD_data <- read.csv("data/GBD_data.csv") %>%
  #mutate(perdist = log(val)) %>%
  group_by(countrycode, Age) %>%
  summarise(meanval = mean(val),
            upperval = mean(upper),
            lowerval = mean(lower)) %>%
  group_by(countrycode) %>%
  mutate(countmeanval = sum(meanval),
         countupperval = sum(upperval),
         countlowerval = sum(lowerval),
         perdist = (meanval / countmeanval)) %>%
  rename(CNTRY = countrycode)

@@

# Table 1
This chunk of code reproduces the data inside Table 1 in the main document.

Note: the output from this table (the object testtbl) is formatted for latex rather than markdown. To completely reproduce that table, the output from testtbl must be pasted into a latex compiler and all forward slashes ("/") must be converted to backward slashes ("\").
```{r Table1}

>>>>

---
title: "Untitled"
author: "Mathew E. Hauer"
date: "April 5, 2018"
output: pdf_document
---

@@

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

@@

```{r Figure1}

@@

```{r Figure2}

@@

```{r suppfig}
GBD_data <- read.csv("data/GBD_data.csv") %>%
  #mutate(perdist = log(val)) %>%
  group_by(countrycode, Age) %>%
  summarise(meanval = mean(val),
            upperval = mean(upper),
            lowerval = mean(lower)) %>%
  group_by(countrycode) %>%
  mutate(countmeanval = sum(meanval),
         countupperval = sum(upperval),
         countlowerval = sum(lowerval),
         perdist = (meanval / countmeanval)) %>%
  rename(CNTRY = countrycode)

@@

```{r Table1}
